\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-13 paper.tex
Reconstructed the section to lift the model into high dimensions

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, boxrule=0pt, enhanced, sharp corners, breakable]}{\end{tcolorbox}}\fi

@@

Consider the high-dimensional data a rectangular matrix \(X\), where
\(X = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}\),
with \(n\) observations in \(p\) dimensions. The objective is to
discover a low-dimensional projection
\(Y = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}\),
represented as an \(n\) × \(d\) matrix, where \(d \ll p\). The reduction
process seeks to remove noise from the original data set while retaining
essential information.

@@

\hypertarget{constructing-the-2d-model}{%
\subsection{Constructing the 2D model}\label{constructing-the-2d-model}}

@@

After computing hexagonal grid configurations, the 2D emdeddings are
binned into hexagon cells. As a result of hexagonal binning , the bin
centroids (\(C_k^{(2)} \equiv (C_{kx}, C_{ky}\))) (see
Figure~\ref{fig-meth} (Step 2)) are obtained.

@@

Creating the model in the high-D space involves a two-step process.
Initially, we identify observations within each hexagonal bin. After
this, we calculate the average of the high-dimensional data within each
hex bin. The resulting values represent the averaged high-dimensional
data points (\(C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p}\))).
Essentially, these averaged high-dimensional data points serve as the
representation of the hex bin centroids in the high-D space.

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, sharp corners, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi

@@

Consider the high-dimensional data a rectangular matrix \(X\), where
\(X = \begin{bmatrix} \textbf{x}_{1} & \textbf{x}_{2} & \cdots & \textbf{x}_{n}\\ \end{bmatrix}^\top\),
with \(n\) observations in \(p\) dimensions. The objective is to
discover a low-dimensional projection
\(Y = \begin{bmatrix} \textbf{y}_{1} & \textbf{y}_{2} & \cdots & \textbf{y}_{n}\\ \end{bmatrix}^\top\),
represented as an \(n\) × \(d\) matrix, where \(d \ll p\). The reduction
process seeks to remove noise from the original data set while retaining
essential information.

@@

\hypertarget{sec-construct2d}{%
\subsection{Constructing the 2D model}\label{sec-construct2d}}

@@

After computing hexagonal grid configurations, the 2D emdeddings are
binned into hexagon cells. As a result of hexagonal binning , the bin
centroids (\(C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2}\)) (see
Figure~\ref{fig-meth} Step 2) are obtained.

@@

\hypertarget{lifting-the-triangular-mesh-points-into-high-dimensions}{%
\subsubsection{Lifting the triangular mesh points into high
dimensions}\label{lifting-the-triangular-mesh-points-into-high-dimensions}}

\textbf{Step1: Cluster 2D points to hexagons}

Expanding upon the information regarding hexagonal binning discussed in
Step 1 of Section~\ref{sec-construct2d}, the primary objective in this
step is to determine the 2D embedding points associated with each
hexagon. As the initial process of hexagonal binning, the 2D points are
clustered into their respective hexagonal bins. By mapping this
information with the hexagonal bin centroids
(\(C_k^{(2)} \equiv (C_{ky_1}, C_{ky_2})\)) that obtained in Step 2 of
the 2D model building (see Section~\ref{sec-construct2d}), we can find
which 2D points are assigned to which data set. (see
Figure~\ref{fig-wkhighD} (a)).

\textbf{Step2: Cluster high dimensional points to hexagons}

Following the step 1, the main focus in this step to determine the
corresponding high dimensional points for each hexagon. Every 2D
embedding point serves as a projection of a data point belonging to high
dimensional space. By using this mapping between the high dimensional
data and their corresponding projections, the high dimensional points
allocated to each hexagons are determined (see video linked in
Figure~\ref{fig-wkhighD}).

\textbf{Step3: Compute the mean within hexagon}

Having identified the high-dimensional points associated with hexagons,
the final step involves computing the mean within each hexagonal bin.
This implies calculating the average of the high-dimensional data points
located within each hex bin. These averaged high-dimensional data
points, denoted as \(C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})\), serve
as the representative coordinates for the hex bin centroids within the
expansive high-dimensional space (see video linked in
Figure~\ref{fig-wkhighD}).

\hypertarget{lifting-the-2d-triangular-mesh-into-high-dimensions}{%
\subsubsection{Lifting the 2D triangular mesh into high
dimensions}\label{lifting-the-2d-triangular-mesh-into-high-dimensions}}

Based on the insights gained in Step 3 of Section~\ref{sec-construct2d},
where connected edges in the 2D triangular mesh were identified, this
step involves lifting these connections into the high dimensions. Having
the mappings of all 2D triangular mesh points into high dimensions, the
points connected in 2D are also connected in high dimensional space (see
video linked in Figure~\ref{fig-wkhighD}).
